---
title: "--- Day 17: Spinlock ---"
author: "annakrystalli"
date: 2017-12-17
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r knitr-opts-chunk, include=FALSE}
# Update knitr chunk options
# https://yihui.name/knitr/options/#chunk-options
knitr::opts_chunk$set(
  comment = NA,
  fig.align = "center",
  tidy = FALSE,
  fig.path = paste0("figure/", knitr::current_input(), "/")
)
```

```{r last-updated, echo=FALSE, results='asis', warning=FALSE}
# Insert the date the file was last updated
cat(sprintf("**Last updated:** %s", Sys.Date()))
```

```{r code-version, echo=FALSE, results='asis'}
# Insert the code version (Git commit SHA1) if Git repository exists and R
# package git2r is installed
if(requireNamespace("git2r", quietly = TRUE)) {
  if(git2r::in_repository()) {
    code_version <- substr(git2r::commits()[[1]]@sha, 1, 7)
  } else {
    code_version <- "Unavailable. Initialize Git repository to enable."
  }
} else {
  code_version <- "Unavailable. Install git2r package to enable."
}
cat(sprintf("**Code version:** %s", code_version))
rm(code_version)
```


> [***See more puzzles***](http://annakrystalli.me/advent_of_code/)

[**Advent of Code**](https://adventofcode.com/2017/)


## Session information

<!-- Insert the session information into the document -->
```{r session-info}
sessionInfo()
```


## Brief

<!-- Insert Part 1 of the puzzle brief here -->

Suddenly, whirling in the distance, you notice what looks like a massive, pixelated hurricane: a deadly spinlock. This spinlock isn't just consuming computing power, but memory, too; vast, digital mountains are being ripped from the ground and consumed by the vortex.

If you don't move quickly, fixing that printer will be the least of your problems.

This spinlock's algorithm is simple but efficient, quickly consuming everything in its path. It starts with a circular buffer containing only the value 0, which it marks as the current position. It then steps forward through the circular buffer some number of steps (your puzzle input) before inserting the first new value, 1, after the value it stopped on. The inserted value becomes the current position. Then, it steps forward from there the same number of steps, and wherever it stops, inserts after it the second new value, 2, and uses that as the new current position again.

It repeats this process of stepping forward, inserting a new value, and using the location of the inserted value as the new current position a total of 2017 times, inserting 2017 as its final operation, and ending with a total of 2018 values (including 0) in the circular buffer.

For example, if the spinlock were to step 3 times per insert, the circular buffer would begin to evolve like this (using parentheses to mark the current position after each iteration of the algorithm):

`(0)`, the initial state before any insertions.
`0 (1):` the spinlock steps forward three times (0, 0, 0), and then inserts the first value, 1, after it. 1 becomes the current position.
`0 (2) 1:` the spinlock steps forward three times (0, 1, 0), and then inserts the second value, 2, after it. 2 becomes the current position.
`0  2 (3) 1:` the spinlock steps forward three times (1, 0, 2), and then inserts the third value, 3, after it. 3 becomes the current position.
And so on:

```
0  2 (4) 3  1
0 (5) 2  4  3  1
0  5  2  4  3 (6) 1
0  5 (7) 2  4  3  6  1
0  5  7  2  4  3 (8) 6  1
0 (9) 5  7  2  4  3  8  6  1
```

Eventually, after 2017 insertions, the section of the circular buffer near the last insertion looks like this:

```1512  1134  151 (2017) 638  1513  851```

Perhaps, if you can identify the value that will ultimately be after the last value written (2017), you can short-circuit the spinlock. In this example, that would be 638.

What is the value after 2017 in your completed circular buffer?



# Let's go

### Packages & functions
```{r, message = F}
library(tidyverse)
library(testthat)
library(aocodeR)
library(hash)
```


## Input

<!-- Supply day. cookie_path defaults to path in my project -->
```{r}
input <- aoc_get_input(day = 17, cookie_path = paste0(rprojroot::find_rstudio_root_file(),
                                                 "/secrets/session_cookie.txt")) %>% as.numeric
input
```

## Functions
```{r}

spin_lock <- function(end = 2017, steps = input) {
    cp <- 1
    circ_buf <- 0
    for(i in 1:end){
        np <- (cp - 1 + steps) %% length(circ_buf) + 1
         circ_buf <- append(circ_buf, i, np)
        cp <- np + 1
    }
    list(cp = cp, circ_buf = circ_buf)
}

next_value <- function(x) {
   x[[2]][x[[1]] + 1]
}

```


```{r}

end <- 5000
t0 <- Sys.time()
t <-spin_lock(end, input)
(Sys.time() - t0) * 50000000/end
```

Time taken per iteration increases with the size of `end` (ie the size of the final vector)


## Test
```{r}
expect_equal(spin_lock(9, 3), 
             list(cp = 2, 
                  circ_buf = c(0, 9, 5,  7,  2,  4,  3,  8,  6,  1))
             )
```

## deploy

`
```{r}
spin_lock(2017, input) %>% next_value
```


## Success!


# ---- Part 2 ----
# `r emojifont::emoji('heavy_exclamation_mark')` HELP NEEDED `r emojifont::emoji('heavy_exclamation_mark')`
#### Too slow, I guess because of the expanding size of the vector.

> **Any feedback? let me know [here](https://github.com/annakrystalli/advent_of_code/issues/4)!**

***

## Brief
<!-- Insert Part 2 of the puzzle brief here -->


The spinlock does not short-circuit. Instead, it gets more angry. At least, you assume that's what happened; it's spinning significantly faster than it was a moment ago.

You have good news and bad news.

The good news is that you have improved calculations for how to stop the spinlock. They indicate that you actually need to identify the value after 0 in the current state of the circular buffer.

The bad news is that while you were determining this, the spinlock has just finished inserting its fifty millionth value (50000000).

What is the value after 0 the moment 50000000 is inserted?

# Let's go

## Test
```{r}
```

## deploy

```{r}
out <- spin_lock(50000000, input)
```


